---
title: "Descriptive Statistics"
output: html_document
---

```{r, echo=FALSE}
loadPath = "C:\\Users\\Sarah\\Google Drive\\PROJECT\\"

```

```{r,echo=FALSE}
e.df = read.csv(file=paste(loadPath, "msci eafe.csv", sep=""), 
                    header=TRUE, stringsAsFactors=FALSE)
M.df = read.csv(file=paste(loadPath, "WILL5000INDFC.csv", sep=""), 
                                 header=TRUE, stringsAsFactors=FALSE)
SG.df = read.csv(file=paste(loadPath, "small cap growth 2000.csv", sep=""), 
                    header=TRUE, stringsAsFactors=FALSE)
SV.df = read.csv(file=paste(loadPath, "small cap value 2000.csv", sep=""), 
                    header=TRUE, stringsAsFactors=FALSE)
LG.df = read.csv(file=paste(loadPath, "russell 1000 growth from march 2000.csv", sep=""), 
                    header=TRUE, stringsAsFactors=FALSE)
LV.df = read.csv(file=paste(loadPath, "russell 1000 value from march 2000.csv", sep=""), 
                    header=TRUE, stringsAsFactors=FALSE)
```

```{r,echo=FALSE}
lg.df<-na.omit(LG.df)
lv.df<-na.omit(LV.df)
sv.df<-na.omit(SV.df)
sg.df<-na.omit(SG.df)
m.df<-na.omit(M.df)
```
```{r,echo=FALSE}
rownames(lg.df) = lg.df$Date
rownames(lv.df) = lv.df$Date
rownames(sv.df) = sv.df$Date
rownames(sg.df) = sg.df$Date
rownames(m.df) = m.df$DATE
rownames(e.df) = e.df$Date
```

```{r, echo=FALSE}
m.df = m.df[, "Adj.Close", drop=FALSE]
e.df = e.df[, "Adj.Close", drop=FALSE]
sg.df = sg.df[, "Adj.Close", drop=FALSE]
sv.df = sv.df[, "Adj.Close", drop=FALSE]
lv.df = lv.df[, "Adj.Close", drop=FALSE]
lg.df = lg.df[, "Adj.Close", drop=FALSE]

```
``` {r, echo=FALSE}

library(PerformanceAnalytics)
library(timeSeries)

```

```{r,echo=FALSE}
dates.e = as.yearmon(rownames(e.df), format="%m/%d/%Y")
dates.sg = as.yearmon(rownames(sg.df), format="%m/%d/%Y")
dates.sv = as.yearmon(rownames(sv.df), format="%m/%d/%Y")
dates.lg = as.yearmon(rownames(lg.df), format="%m/%d/%Y")
dates.lv = as.yearmon(rownames(lv.df), format="%m/%d/%Y")
dates.m = as.yearmon(rownames(m.df), format="%m/%d/%Y")
```
```{r,echo=FALSE}
sg.z = zoo(x=sg.df$Adj.Close, order.by=dates.sg)
sv.z = zoo(x=sv.df$Adj.Close, order.by=dates.sv)
lg.z = zoo(x=lg.df$Adj.Close, order.by=dates.lg)
lv.z = zoo(x=lv.df$Adj.Close, order.by=dates.lv)
m.z = zoo(x=m.df$Adj.Close, order.by=dates.m)
e.z = zoo(x=e.df$Adj.Close, order.by=dates.e)
```
##**Calculating Compound Return**

```{r,echo=FALSE}

ec.z = (CalculateReturns(e.z, method="compound"))[-1]
sgc.z =CalculateReturns(sg.z, method="simple")[-1]
mc.z = CalculateReturns(m.z, method="compound")[-1]
svc.z = CalculateReturns(sv.z, method="compound")[-1]
lvc.z = CalculateReturns(lv.z, method="compound")[-1]
lgc.z = CalculateReturns(lg.z, method="compound")[-1]
```

```{r,echo=FALSE}
ac.z=merge(sgc.z,ec.z,svc.z,lvc.z,lgc.z,mc.z)
ab.z=merge(ec.z,mc.z)

chart.TimeSeries(ab.z,lwd=2,main="Returns on US vs European Market",legend.loc="bottomright")
chart.CumReturns(ac.z, lwd=2, main="Growth of $1", legend.loc="topleft")
e.m<-as.matrix(ec.z)
rownames(e.m)<-rownames(e.df)[-1]
sv.m<-as.matrix(svc.z)
rownames(sv.m)<-rownames(sv.df)[-1]
sg.m<-as.matrix(sgc.z)
rownames(sg.m)<-rownames(sg.df)[-1]
lv.m<-as.matrix(lvc.z)
rownames(lv.m)<-rownames(lv.df)[-1]
lg.m<-as.matrix(lgc.z)
rownames(lg.m)<-rownames(lg.df)[-1]
m.m<-m.m<-as.matrix(mc.z)
rownames(m.m)<-rownames(m.df)[-1]
a.m<-as.matrix(ac.z)[,-1]
rownames(a.m)<-rownames(m.df)[-1]

```
##**Examining the Frequency Distribution**

```{r,echo=FALSE}

chart.Histogram((sv.m),methods=c("add.density","add.risk","add.rug","add.qqplot","add.centered"))
chart.Histogram((lg.m),methods=c("add.density","add.risk","add.rug","add.qqplot","add.centered"))
```
###**Charting Asset Classes with same Histogram Breaks to Facilitate Comparison**
```{r,echo=FALSE}
m.hist = hist(m.m,plot=F,breaks=10)
class(m.hist)
names(m.hist)      

hist(m.m,main="US", col="slateblue1", xlab="returns", breaks=m.hist$breaks)
hist(e.m,main="Europe", col="slateblue1", xlab="returns",breaks=m.hist$breaks)
hist(lg.m,main="US", col="slateblue1", xlab="returns", breaks=m.hist$breaks)
hist(lv.m,main="Europe", col="slateblue1", xlab="returns",breaks=m.hist$breaks)

```

###Various Statistics for Each Market 
US Equity 
```{r}
hist(m.m,main="US Equity", col="slateblue1",
     probability=T, ylim=c(0,18))
market.density = density(m.m[-1])
points(market.density,type="l", col="orange", lwd=2)
median(m.m)
quantile(m.m,probs=0.75) - quantile(m.m,probs=0.25)
mean(m.m)
var(m.m)
sd(m.m)
skewness(m.m)
kurtosis(m.m)
min(m.m)
max(m.m)
```
Europe Equity
```{r}
hist(e.m,main="Europe Equity", col="slateblue1",
     probability=T, ylim=c(0,18))
market.density = density(e.m[-1])
points(market.density,type="l", col="orange", lwd=2)
median(e.m)
quantile(e.m,probs=0.75) - quantile(e.m,probs=0.25)
mean(e.m)
var(e.m)
sd(e.m)
skewness(e.m)
kurtosis(e.m)
min(e.m)
max(e.m)
```


## **Box Plots and Scatter Plots**

```{r,echo=FALSE}
boxplot.matrix(a.m,names=c("Euro","SV","LV","LG","US"),outchar=T, col="slateblue1",
               main="Comparison of return distributions",ylab="monthly return")
boxplot(sg.m)

plot(m.m,e.m,main="Monthly cc returns on Dev and Em", col="slateblue1")
abline(h=mean(e.m))  # horizontal line at SP500 mean
abline(v=mean(m.m))    # vertical line at MSFT mean

pairs(a.m, col="slateblue1") 
```

*most relationships are between both types of value stocks, and between large stocksThe MSCI has poor correlation with all.* 


## **Comparing the Market Return with that of an IDD Gaussian Distribution with the same Mean and SD as US Equity**

```{r}

set.seed(123)
gwn = rnorm(length(m.m),mean=mean(m.m),sd=sd(m.m))
```
```{r,echo=FALSE}
par(mfrow=c(2,1))
ts.plot(m.m,main="US Equity", lwd=2, col="blue")
abline(h=0)
ts.plot(gwn,main="Gaussian data with same mean and sd as US Equity", lwd=2, col="blue")
abline(h=0)
par(mfrow=c(1,1))

z1 = scale(gwn)          # standardize to have mean zero and sd 1
n1 = length(gwn)
F.hat = 1:n1/n1        	# empirical cdf
x1 = sort(z1)				# sort from smallest to largest
y1 = pnorm(x1)			# compute standard normal cdf at x

plot(x1,y1,main="Empirical CDF vs. Normal CDF for Gaussian data",
     type="l",lwd=2,xlab="standardized gwn",ylab="CDF")
points(x1,F.hat, type="s", lty=1, lwd=3, col="orange")

legend(x="topleft",legend=c("Normal CDF","Empirical CDF"),
       lty=c(1,1), lwd=2, col=c("black","orange"))

```



```{r,echo=FALSE}

quantile(m.m,probs=c(0.01,0.05),na.rm=T)

qnorm(p=c(0.01,0.05), mean=mean(m.m), sd=sd(m.m))

qqnorm(gwn, main="Gaussian data")
qqline(gwn)
qqnorm(m.m, main="US")
qqline(m.m)
qqnorm(e.m, main="Europe")
qqline(e.m)
qqnorm(sv.m, main="Large Growth")
qqline(sv.m)
qqnorm(lg.m, main="Large Growth")
qqline(lg.m)
qqnorm(lv.m, main="Large Growth")
qqline(lv.m)
```
#Constant Expected Return Model Calibrated to monthly market return

```{r}
mu = mean(m.m)
sd.e = sd(m.m)
nobs = 170
set.seed(114)
sim.e = rnorm(nobs, mean=0, sd=sd.e)
sim.ret = mu + sim.e
boxplot(cbind(m.m,sim.ret,gwn), names=c("World","CER","GWN"))
```

### **Comparing Market Return with the Random Walk Model, controlling for SD of errors and Mean of return** UNDERSTAND AND FIX 

```{r}
mu = mean(m.m)
sd.e = sd(m.m)
nobs = 170
set.seed(132)
sim.e = rnorm(nobs, mean=0, sd=sd.e)
sim.p = 1 + mu*seq(nobs) + cumsum(sim.e)
sim.P = exp(sim.p)
par(mfrow=c(2,1),mar=rep(2,4))
ts.plot(sim.p, col="blue",lwd=2,
        ylim=c(-2, 4), ylab="log price")
lines( (1+mu*seq(nobs)), lty="dotted", col="black", lwd=2)
lines(cumsum(sim.e), col="orange", lty="dashed", lwd=2)
abline(h=0)
legend(x="topleft",legend=c("p(t)","E[p(t)]","p(t)-E[p(t)]"),
       lty=c("solid","dotted","dashed"), col=c("blue","black","orange"), 
       lwd=2, cex=c(0.75,0.75,0.75))
ts.plot(sim.P, lwd=2, col="blue", ylab="price")

boxplot(cbind(sim.e,m.m), lwd=2, col=c("green","yellow"), names=c("random walk","market"),ylab="price")

```

### Computing means, variances and covariances 
```{r}
muhat.vals = colMeans(a.m)
muhat.vals
sigma2hat.vals = colVars(a.m)
sigma2hat.vals
sigmahat.vals = colSds(a.m)
sigmahat.vals
cov.mat = cov(a.m)
cov.mat
cor.mat = cor(a.m)
cor.mat
covhat.vals = cov.mat[lower.tri(cov.mat)]
rhohat.vals = cor.mat[lower.tri(cor.mat)]
names(covhat.vals) <- names(rhohat.vals) <- c("sve","lve","lge","me","lvsv","lgsv","msv","lvlg","lvm","lgm")
```
### compute estimated standard error for mean
```{r}

nobs = nrow(a.m)
nobs
se.muhat = sigmahat.vals/sqrt(nobs)
se.muhat # in some cases the se of the mean is even bigger than the mean itself! (in LG)

# compute t-ratios
muhat.vals/se.muhat
# compute exact 95% confidence intervals
t.975 = qt(0.975, df=99)
mu.lower = muhat.vals - t.975*se.muhat
mu.upper = muhat.vals + t.975*se.muhat
mu.width = mu.upper - mu.lower
cbind(mu.lower,mu.upper,mu.width)
```

### compute estimated standard error for Variance and Standard Deviation

```{r}
se.sigma2hat = sigma2hat.vals/sqrt(nobs/2)
se.sigmahat = sigmahat.vals/sqrt(2*nobs)
se.sigma2hat
se.sigmahat
# compute approx 95% confidence intervals
sigma2.lower = sigma2hat.vals - 2*se.sigma2hat
sigma2.upper = sigma2hat.vals + 2*se.sigma2hat
sigma2.width = sigma2.upper - sigma2.lower
cbind(sigma2.lower,sigma2.upper,sigma2.width)

sigma.lower = sigmahat.vals - 2*se.sigmahat
sigma.upper = sigmahat.vals + 2*se.sigmahat
sigma.width = sigma.upper - sigma.lower
cbind(sigma.lower,sigma.upper,sigma.width)
```

###compute estimated standard error for Correlation

```{r}
se.rhohat = (1-rhohat.vals^2)/sqrt(nobs)
se.rhohat

# compute approx 95% confidence intervals
rho.lower = rhohat.vals - 2*se.rhohat
rho.upper = rhohat.vals + 2*se.rhohat
rho.width = rho.upper - rho.lower
cbind(rho.lower,rho.upper,rho.width)

```

# show risk return tradeoffs
```{r}
cex.val = 2
plot(sigmahat.vals, muhat.vals,  ylim=c(0, 0.04), xlim=c(0, 0.20), ylab=expression(mu[p]),
     xlab=expression(sigma[p]), pch=16, col="blue", cex=2.5, cex.lab=1.75)     
text(sigmahat.vals, muhat.vals, labels=names(muhat.vals), pos=4, cex = cex.val)
```

# optimization 

```{r,echo=FALSE}

# portfolio.r
# 
# Functions for portfolio analysis
# to be used in Introduction to Computational Finance & Financial Econometrics
# last updated: August 8, 2012 by Hezky Varon
#               November 7, 2000 by Eric Zivot
#               Oct 15, 2003 by Tim Hesterberg
#               November 18, 2003 by Eric Zivot
#               November 9, 2004 by Eric Zivot
#    	            November 9, 2008 by Eric Zivot
#               August 11, 2011 by Eric Zivot
#
# Functions:
#	1. efficient.portfolio			compute minimum variance portfolio
#							                subject to target return
#	2. globalMin.portfolio			compute global minimum variance portfolio
#	3. tangency.portfolio			  compute tangency portfolio
#	4. efficient.frontier			  compute Markowitz bullet
#	5. getPortfolio					    create portfolio object

stopifnot("package:quadprog" %in% search()  ||  require("quadprog",quietly = TRUE) )

getPortfolio <-
function(er, cov.mat, weights)
{
	# contruct portfolio object
	#
	# inputs:
	# er				   N x 1 vector of expected returns
	# cov.mat  		 N x N covariance matrix of returns
	# weights			 N x 1 vector of portfolio weights
	#
	# output is portfolio object with the following elements
	# call				original function call
	# er				  portfolio expected return
	# sd				  portfolio standard deviation
	# weights			N x 1 vector of portfolio weights
	#
	call <- match.call()
	
	#
	# check for valid inputs
	#
	asset.names <- names(er)
	weights <- as.vector(weights)
	names(weights) = names(er)
  er <- as.vector(er)					# assign names if none exist
	if(length(er) != length(weights))
		stop("dimensions of er and weights do not match")
 	cov.mat <- as.matrix(cov.mat)
	if(length(er) != nrow(cov.mat))
		stop("dimensions of er and cov.mat do not match")
	if(any(diag(chol(cov.mat)) <= 0))
		stop("Covariance matrix not positive definite")
		
	#
	# create portfolio
	#
	er.port <- crossprod(er,weights)
	sd.port <- sqrt(weights %*% cov.mat %*% weights)
	ans <- list("call" = call,
	      "er" = as.vector(er.port),
	      "sd" = as.vector(sd.port),
	      "weights" = weights) 
	class(ans) <- "portfolio"
	ans
}

efficient.portfolio <-
function(er, cov.mat, target.return, shorts=TRUE)
{
  # compute minimum variance portfolio subject to target return
  #
  # inputs:
  # er					    N x 1 vector of expected returns
  # cov.mat  			  N x N covariance matrix of returns
  # target.return	  scalar, target expected return
  # shorts          logical, allow shorts is TRUE
  #
  # output is portfolio object with the following elements
  # call				    original function call
  # er					    portfolio expected return
  # sd					    portfolio standard deviation
  # weights			    N x 1 vector of portfolio weights
  #
  call <- match.call()

  #
  # check for valid inputs
  #
  asset.names <- names(er)
  er <- as.vector(er)					# assign names if none exist
  N <- length(er)
  cov.mat <- as.matrix(cov.mat)
  if(N != nrow(cov.mat))
    stop("invalid inputs")
  if(any(diag(chol(cov.mat)) <= 0))
    stop("Covariance matrix not positive definite")
  # remark: could use generalized inverse if cov.mat is positive semidefinite

  #
  # compute efficient portfolio
  #
  if(shorts==TRUE){
    ones <- rep(1, N)
    top <- cbind(2*cov.mat, er, ones)
    bot <- cbind(rbind(er, ones), matrix(0,2,2))
    A <- rbind(top, bot)
    b.target <- as.matrix(c(rep(0, N), target.return, 1))
    x <- solve(A, b.target)
    w <- x[1:N]
  } else if(shorts==FALSE){
    Dmat <- 2*cov.mat
    dvec <- rep.int(0, N)
    Amat <- cbind(rep(1,N), er, diag(1,N))
    bvec <- c(1, target.return, rep(0,N))
    result <- solve.QP(Dmat=Dmat,dvec=dvec,Amat=Amat,bvec=bvec,meq=2)
    w <- round(result$solution, 6)
  } else {
    stop("shorts needs to be logical. For no-shorts, shorts=FALSE.")
  }

  #
  # compute portfolio expected returns and variance
  #
  names(w) <- asset.names
  er.port <- crossprod(er,w)
  sd.port <- sqrt(w %*% cov.mat %*% w)
  ans <- list("call" = call,
	      "er" = as.vector(er.port),
	      "sd" = as.vector(sd.port),
	      "weights" = w) 
  class(ans) <- "portfolio"
  ans
}

globalMin.portfolio <-
function(er, cov.mat, shorts=TRUE)
{
  # Compute global minimum variance portfolio
  #
  # inputs:
  # er				N x 1 vector of expected returns
  # cov.mat		N x N return covariance matrix
  # shorts          logical, allow shorts is TRUE
  #
  # output is portfolio object with the following elements
  # call			original function call
  # er				portfolio expected return
  # sd				portfolio standard deviation
  # weights		N x 1 vector of portfolio weights
  call <- match.call()

  #
  # check for valid inputs
  #
  asset.names <- names(er)
  er <- as.vector(er)					# assign names if none exist
  cov.mat <- as.matrix(cov.mat)
  N <- length(er)
  if(N != nrow(cov.mat))
    stop("invalid inputs")
  if(any(diag(chol(cov.mat)) <= 0))
    stop("Covariance matrix not positive definite")
  # remark: could use generalized inverse if cov.mat is positive semi-definite

  #
  # compute global minimum portfolio
  #
  if(shorts==TRUE){
    cov.mat.inv <- solve(cov.mat)
    one.vec <- rep(1,N)
    w.gmin <- rowSums(cov.mat.inv) / sum(cov.mat.inv)
    w.gmin <- as.vector(w.gmin)
  } else if(shorts==FALSE){
    Dmat <- 2*cov.mat
    dvec <- rep.int(0, N)
    Amat <- cbind(rep(1,N), diag(1,N))
    bvec <- c(1, rep(0,N))
    result <- solve.QP(Dmat=Dmat,dvec=dvec,Amat=Amat,bvec=bvec,meq=1)
    w.gmin <- round(result$solution, 6)
  } else {
    stop("shorts needs to be logical. For no-shorts, shorts=FALSE.")
  }
 
  names(w.gmin) <- asset.names
  er.gmin <- crossprod(w.gmin,er)
  sd.gmin <- sqrt(t(w.gmin) %*% cov.mat %*% w.gmin)
  gmin.port <- list("call" = call,
		    "er" = as.vector(er.gmin),
		    "sd" = as.vector(sd.gmin),
		    "weights" = w.gmin)
  class(gmin.port) <- "portfolio"
  gmin.port
}


tangency.portfolio <- 
function(er,cov.mat,risk.free, shorts=TRUE)
{
  # compute tangency portfolio
  #
  # inputs:
  # er				   N x 1 vector of expected returns
  # cov.mat		   N x N return covariance matrix
  # risk.free		 scalar, risk-free rate
  # shorts          logical, allow shorts is TRUE
  #
  # output is portfolio object with the following elements
  # call			  captures function call
  # er				  portfolio expected return
  # sd				  portfolio standard deviation
  # weights		 N x 1 vector of portfolio weights
  call <- match.call()

  #
  # check for valid inputs
  #
  asset.names <- names(er)
  if(risk.free < 0)
    stop("Risk-free rate must be positive")
  er <- as.vector(er)
  cov.mat <- as.matrix(cov.mat)
  N <- length(er)
  if(N != nrow(cov.mat))
    stop("invalid inputs")
  if(any(diag(chol(cov.mat)) <= 0))
    stop("Covariance matrix not positive definite")
  # remark: could use generalized inverse if cov.mat is positive semi-definite

  #
  # compute global minimum variance portfolio
  #
  gmin.port <- globalMin.portfolio(er, cov.mat, shorts=shorts)
  if(gmin.port$er < risk.free)
    stop("Risk-free rate greater than avg return on global minimum variance portfolio")

  # 
  # compute tangency portfolio
  #
  if(shorts==TRUE){
    cov.mat.inv <- solve(cov.mat)
    w.t <- cov.mat.inv %*% (er - risk.free) # tangency portfolio
    w.t <- as.vector(w.t/sum(w.t))	# normalize weights
  } else if(shorts==FALSE){
    Dmat <- 2*cov.mat
    dvec <- rep.int(0, N)
    er.excess <- er - risk.free
    Amat <- cbind(er.excess, diag(1,N))
    bvec <- c(1, rep(0,N))
    result <- solve.QP(Dmat=Dmat,dvec=dvec,Amat=Amat,bvec=bvec,meq=1)
    w.t <- round(result$solution/sum(result$solution), 6)
  } else {
    stop("Shorts needs to be logical. For no-shorts, shorts=FALSE.")
  }
    
  names(w.t) <- asset.names
  er.t <- crossprod(w.t,er)
  sd.t <- sqrt(t(w.t) %*% cov.mat %*% w.t)
  tan.port <- list("call" = call,
		   "er" = as.vector(er.t),
		   "sd" = as.vector(sd.t),
		   "weights" = w.t)
  class(tan.port) <- "portfolio"
  tan.port
}

efficient.frontier <- 
function(er, cov.mat, nport=20, alpha.min=-0.5, alpha.max=1.5, shorts=TRUE)
{
  # Compute efficient frontier with no short-sales constraints
  #
  # inputs:
  # er			  N x 1 vector of expected returns
  # cov.mat	  N x N return covariance matrix
  # nport		  scalar, number of efficient portfolios to compute
  # shorts          logical, allow shorts is TRUE
  #
  # output is a Markowitz object with the following elements
  # call		  captures function call
  # er			  nport x 1 vector of expected returns on efficient porfolios
  # sd			  nport x 1 vector of std deviations on efficient portfolios
  # weights 	nport x N matrix of weights on efficient portfolios 
  call <- match.call()

  #
  # check for valid inputs
  #
  asset.names <- names(er)
  er <- as.vector(er)
  N <- length(er)
  cov.mat <- as.matrix(cov.mat)
  if(N != nrow(cov.mat))
    stop("invalid inputs")
  if(any(diag(chol(cov.mat)) <= 0))
    stop("Covariance matrix not positive definite")

  #
  # create portfolio names
  #
  port.names <- rep("port",nport)
  ns <- seq(1,nport)
  port.names <- paste(port.names,ns)

  #
  # compute global minimum variance portfolio
  #
  cov.mat.inv <- solve(cov.mat)
  one.vec <- rep(1, N)
  port.gmin <- globalMin.portfolio(er, cov.mat, shorts)
  w.gmin <- port.gmin$weights

  if(shorts==TRUE){
    # compute efficient frontier as convex combinations of two efficient portfolios
    # 1st efficient port: global min var portfolio
    # 2nd efficient port: min var port with ER = max of ER for all assets
    er.max <- max(er)
    port.max <- efficient.portfolio(er,cov.mat,er.max)
    w.max <- port.max$weights    
    a <- seq(from=alpha.min,to=alpha.max,length=nport)			# convex combinations
    we.mat <- a %o% w.gmin + (1-a) %o% w.max	# rows are efficient portfolios
    er.e <- we.mat %*% er							# expected returns of efficient portfolios
    er.e <- as.vector(er.e)
  } else if(shorts==FALSE){
    we.mat <- matrix(0, nrow=nport, ncol=N)
    we.mat[1,] <- w.gmin
    we.mat[nport, which.max(er)] <- 1
    er.e <- as.vector(seq(from=port.gmin$er, to=max(er), length=nport))
    for(i in 2:(nport-1)) 
      we.mat[i,] <- efficient.portfolio(er, cov.mat, er.e[i], shorts)$weights
  } else {
    stop("shorts needs to be logical. For no-shorts, shorts=FALSE.")
  }
  
  names(er.e) <- port.names
  cov.e <- we.mat %*% cov.mat %*% t(we.mat) # cov mat of efficient portfolios
  sd.e <- sqrt(diag(cov.e))					# std devs of efficient portfolios
  sd.e <- as.vector(sd.e)
  names(sd.e) <- port.names
  dimnames(we.mat) <- list(port.names,asset.names)

  # 
  # summarize results
  #
  ans <- list("call" = call,
	      "er" = er.e,
	      "sd" = sd.e,
	      "weights" = we.mat)
  class(ans) <- "Markowitz"
  ans
}

#
# print method for portfolio object
print.portfolio <- function(x, ...)
{
  cat("Call:\n")
  print(x$call, ...)
  cat("\nPortfolio expected return:    ", format(x$er, ...), "\n")
  cat("Portfolio standard deviation: ", format(x$sd, ...), "\n")
  cat("Portfolio weights:\n")
  print(round(x$weights,4), ...)
  invisible(x)
}

#
# summary method for portfolio object
summary.portfolio <- function(object, risk.free=NULL, ...)
# risk.free			risk-free rate. If not null then
#				compute and print Sharpe ratio
# 
{
  cat("Call:\n")
  print(object$call)
  cat("\nPortfolio expected return:    ", format(object$er, ...), "\n")
  cat(  "Portfolio standard deviation: ", format(object$sd, ...), "\n")
  if(!is.null(risk.free)) {
    SharpeRatio <- (object$er - risk.free)/object$sd
    cat("Portfolio Sharpe Ratio:       ", format(SharpeRatio), "\n")
  }
  cat("Portfolio weights:\n")
  print(round(object$weights,4), ...)
  invisible(object)
}
# hard-coded 4 digits; prefer to let user control, via ... or other argument

#
# plot method for portfolio object
plot.portfolio <- function(object, ...)
{
  asset.names <- names(object$weights)
  barplot(object$weights, names=asset.names,
	  xlab="Assets", ylab="Weight", main="Portfolio Weights", ...)
  invisible()
}

#
# print method for Markowitz object
print.Markowitz <- function(x, ...)
{
  cat("Call:\n")
  print(x$call)
  xx <- rbind(x$er,x$sd)
  dimnames(xx)[[1]] <- c("ER","SD")
  cat("\nFrontier portfolios' expected returns and standard deviations\n")
  print(round(xx,4), ...)
  invisible(x)
}
# hard-coded 4, should let user control

#
# summary method for Markowitz object
summary.Markowitz <- function(object, risk.free=NULL)
{
  call <- object$call
  asset.names <- colnames(object$weights)
  port.names <- rownames(object$weights)
  if(!is.null(risk.free)) {
    # compute efficient portfolios with a risk-free asset
    nport <- length(object$er)
    sd.max <- object$sd[1]
    sd.e <- seq(from=0,to=sd.max,length=nport)	
    names(sd.e) <- port.names

    #
    # get original er and cov.mat data from call 
    er <- eval(object$call$er)
    cov.mat <- eval(object$call$cov.mat)

    #
    # compute tangency portfolio
    tan.port <- tangency.portfolio(er,cov.mat,risk.free)
    x.t <- sd.e/tan.port$sd		# weights in tangency port
    rf <- 1 - x.t			# weights in t-bills
    er.e <- risk.free + x.t*(tan.port$er - risk.free)
    names(er.e) <- port.names
    we.mat <- x.t %o% tan.port$weights	# rows are efficient portfolios
    dimnames(we.mat) <- list(port.names, asset.names)
    we.mat <- cbind(rf,we.mat) 
  }
  else {
    er.e <- object$er
    sd.e <- object$sd
    we.mat <- object$weights
  }
  ans <- list("call" = call,
	      "er"=er.e,
	      "sd"=sd.e,
	      "weights"=we.mat)
  class(ans) <- "summary.Markowitz"	
  ans
}

print.summary.Markowitz <- function(x, ...)
{
	xx <- rbind(x$er,x$sd)
	port.names <- names(x$er)
	asset.names <- colnames(x$weights)
	dimnames(xx)[[1]] <- c("ER","SD")
	cat("Frontier portfolios' expected returns and standard deviations\n")
	print(round(xx,4), ...)
	cat("\nPortfolio weights:\n")
	print(round(x$weights,4), ...)
	invisible(x)
}
# hard-coded 4, should let user control

#
# plot efficient frontier
#
# things to add: plot original assets with names
# tangency portfolio
# global min portfolio
# risk free asset and line connecting rf to tangency portfolio
#
plot.Markowitz <- function(object, plot.assets=FALSE, ...)
# plot.assets		logical. If true then plot asset sd and er
{
  if (!plot.assets) {
     y.lim=c(0,max(object$er))
     x.lim=c(0,max(object$sd))
     plot(object$sd,object$er,type="b",xlim=x.lim, ylim=y.lim,
          xlab="Portfolio SD", ylab="Portfolio ER", 
          main="Efficient Frontier", ...)
     }
  else {
	  call = object$call
	  mu.vals = eval(call$er)
	  sd.vals = sqrt( diag( eval(call$cov.mat) ) )
	  y.lim = range(c(0,mu.vals,object$er))
	  x.lim = range(c(0,sd.vals,object$sd))
	  plot(object$sd,object$er,type="b", xlim=x.lim, ylim=y.lim,
          xlab="Portfolio SD", ylab="Portfolio ER", 
          main="Efficient Frontier", ...)
        text(sd.vals, mu.vals, labels=names(mu.vals))
  }
  invisible()
}


```
```{r}
# Data Needed for Optimization :to adjust return and correlation towards 15-years average
#last 5 years of return
a.m1=a.m[110:170,]
muhat=colMeans(a.m1[,-5])
a=(a.m1[,-5])

# Adjusting Cov Matrix and Cov for last 5 years 
covmat=cov(a.m)[-5,-5]
cov.mat1 = cov(a.m1)[-5,-5]
covhat.vals1 = cov.mat1[lower.tri(cov.mat1)]
covhatvals=covmat[lower.tri(covmat)]

# Adjusting Historic Returns to get to Expected Returns 
r.free=0.0011
er=muhat*(2/3)+muhat.vals[-5]*(1/3)
er=er[-5]

#Expected Cov Matrix
ecov=cov.mat1*(2/3)+covmat*(1/3)
names(covhat.vals1)=names(covhatvals)=names(ecov)<- c("sve","lve","lge","lvsv","lgsv","lvlg")


# compute equally weighted portfolio
ew = rep(1,4)/4
equalWeight.portfolio = getPortfolio(er=er,cov.mat=ecov,weights=ew)
class(equalWeight.portfolio)
names(equalWeight.portfolio)
equalWeight.portfolio
summary(equalWeight.portfolio)
plot(equalWeight.portfolio, col="blue")

#
# compute global minimum variance portfolio
gmin.port = globalMin.portfolio(er, ecov)
attributes(gmin.port)
print(gmin.port)
summary(gmin.port, risk.free=r.free)
plot(gmin.port, col="blue")


gmin.port.ns = globalMin.portfolio(er, ecov, shorts=FALSE)
attributes(gmin.port.ns)
print(gmin.port.ns)
summary(gmin.port.ns, risk.free=r.free)
plot(gmin.port.ns, col="blue")

# compute efficient portfolio subject to target return
target.return = er["svc.z"]
e.port.msft = efficient.portfolio(er,ecov, target.return)
e.port.msft
summary(e.port.msft, risk.free=r.free)
plot(e.port.msft, col="blue")

#
# compute efficient portfolio subject to target return with no short sales
target.return = er["svc.z"]
e.port.msft.ns = efficient.portfolio(er, ecov, target.return, shorts=FALSE)
e.port.msft.ns
summary(e.port.msft.ns, risk.free=r.free)
plot(e.port.msft.ns, col="blue")


# compute tangency portfolio
tan.port <- tangency.portfolio(er, ecov, r.free)
tan.port
summary(tan.port, risk.free=r.free)
plot(tan.port, col="blue")

#
# compute tangency portfolio with no short sales
tan.port.ns <- tangency.portfolio(er, ecov, r.free, shorts=FALSE)
tan.port.ns
summary(tan.port.ns, risk.free=r.free)
plot(tan.port.ns, col="blue")

# compute portfolio frontier
ef <- efficient.frontier(er, ecov, alpha.min=-2, 
                         alpha.max=1.5, nport=20)
attributes(ef)
ef

plot(ef)
plot(ef, plot.assets=TRUE, col="blue", pch=16)
points(gmin.port$sd, gmin.port$er, col="green", pch=16, cex=2)
points(tan.port$sd, tan.port$er, col="red", pch=16, cex=2)
text(gmin.port$sd, gmin.port$er, labels="GLOBAL MIN", pos=2)
text(tan.port$sd, tan.port$er, labels="TANGENCY", pos=2)    
sr.tan = (tan.port$er - r.free)/tan.port$sd
abline(a=r.free, b=sr.tan, col="green", lwd=2)


# plot portfolio frontier with tangency portfolio
sd.vals = sqrt(diag(ecov))
mu.vals = er
plot(ef$sd, ef$er, ylim=c(0, max(ef$er)), xlim=c(0, max(ef$sd)),
     xlab="portfolio sd", ylab="portfolio er", main="Efficient Portfolios")
text(sd.vals, mu.vals, labels=names(mu.vals))
abline(a=r.free, b=sr.tan)

#
# compute portfolio frontier with no short sales
ef.ns <- efficient.frontier(er, ecov, alpha.min=0, 
                            alpha.max=1, nport=20, shorts=FALSE)
attributes(ef.ns)
ef.ns
summary(ef.ns)

plot(ef.ns)
plot(ef.ns, plot.assets=TRUE, col="blue", pch=16)
points(gmin.port.ns$sd, gmin.port.ns$er, col="green", pch=16, cex=2)
points(tan.port.ns$sd, tan.port.ns$er, col="red", pch=16, cex=2)
text(gmin.port.ns$sd, gmin.port.ns$er, labels="GLOBAL MIN", pos=2)
text(tan.port.ns$sd, tan.port.ns$er, labels="TANGENCY", pos=2)    
sr.tan.ns = (tan.port.ns$er - r.free)/tan.port.ns$sd
abline(a=r.free, b=sr.tan.ns, col="green", lwd=2)
# plot portfolio frontier with tangency portfolio
sd.vals = sqrt(diag(ecov))
mu.vals = er
plot(ef.ns$sd, ef.ns$er, ylim=c(0, max(ef.ns$er)), xlim=c(0, max(ef.ns$sd)),
     xlab="portfolio sd", ylab="portfolio er", main="Efficient Portfolios")
text(sd.vals, mu.vals, labels=names(mu.vals))
abline(a=r.free, b=sr.tan.ns)

# BOOTSTRAPPING
options(digits=4, width=70)
library(boot) # note: don't actually use the boot library for examples
library(PerformanceAnalytics)

n.obs=nrow(a)
n.boot = 999
mu.boot = matrix(0, n.boot, ncol(a))
sd.boot = matrix(0, n.boot, ncol(a))
colnames(mu.boot) = colnames(sd.boot) = colnames(a)

set.seed(144)
for (i in 1:n.boot) {
    boot.idx = sample(n.obs, replace=TRUE)
    ret.boot = a[boot.idx, ] 
    mu.boot[i, ] = colMeans(ret.boot)
    sd.boot[i, ] = apply(ret.boot, 2, sd) 
}

plot(sd.boot[, "lgc.z"], mu.boot[, "lgc.z"], col="black", pch=16,
     ylim=c(-0.03, 0.07), xlim=c(0, 0.20),
     ylab=expression(mu[p]),
     xlab=expression(sigma[p]), cex.lab=1.5)
points(sd.vals["lgc.z"], mu.vals["lgc.z"], pch=16, col="blue", cex=2.5)
text(sd.vals["lgc.z"], mu.vals["lgc.z"], labels="lgc.z", pos=4, cex = 2)

# plot boeing
points(sd.boot[, "svc.z"], mu.boot[, "svc.z"], col="blue", pch=16)
points(sd.vals["svc.z"], mu.vals["svc.z"], pch=16, col="blue", cex=2.5)
text(sd.vals["svc.z"], mu.vals["svc.z"], labels="Small Value", pos=4, cex = 2)
# plot nordstrom
points(sd.boot[, "ec.z"], mu.boot[, "ec.z"], col="green", pch=16)
points(sd.vals["ec.z"], mu.vals["ec.z"], pch=16, col="green", cex=2.5)
text(sd.vals["ec.z"], mu.vals["ec.z"], labels="Europe", pos=4, cex = 2)
# plot microsoft
points(sd.boot[, "lvc.z"], mu.boot[, "lvc.z"], col="red", pch=16)
points(sd.vals["lvc.z"], mu.vals["lvc.z"], pch=16, col="red", cex=2.5)
text(sd.vals["lvc.z"], mu.vals["lvc.z"], labels="Large Value", pos=4, cex = 2)

# show global minimum variance portfolio
gmin.port = globalMin.portfolio(mu.vals, ecov)
gmin.port

# show risk return tradeoffs with global min
plot(sd.vals, mu.vals,  ylim=c(0, 0.04), xlim=c(0, 0.20), ylab=expression(mu[p]),
     xlab=expression(sigma[p]), pch=16, col="blue", cex=2.5, cex.lab=1.75)     
text(sd.vals, mu.vals, labels=names(mu.vals), pos=4, cex = cex.val)
points(gmin.port$sd, gmin.port$er, pch=16, cex=2.5, col="green")
text(gmin.port$sd, gmin.port$er, labels="Global Min", pos=2, cex = cex.val)

# bootstrap global min portfolio
mu.gmin.boot = matrix(0, n.boot, 1)
sd.gmin.boot = matrix(0, n.boot, 1)
w.gmin.boot = matrix(0, n.boot, 4)
colnames(mu.gmin.boot) = colnames(sd.gmin.boot) = "global.min"
colnames(w.gmin.boot) = names(mu.vals)

set.seed(144)
for (i in 1:n.boot) {
    boot.idx = sample(n.obs, replace=TRUE)
    ret.boot = a[boot.idx, ] 
    mu.boot = colMeans(ret.boot)
    cov.boot = cov(ret.boot) 
    gmin.boot = globalMin.portfolio(mu.boot, cov.boot)
    mu.gmin.boot[i, ] = gmin.boot$er
    sd.gmin.boot[i, ] = gmin.boot$sd
    w.gmin.boot[i, ] = gmin.boot$weights
}

plot(sd.vals, mu.vals,  ylim=c(-0.01, 0.04), xlim=c(0, 0.20), ylab=expression(mu[p]),
     xlab=expression(sigma[p]), pch=16, col="blue", cex=2.5, cex.lab=1.75)     
abline(h=0, v=0)
text(sd.vals, mu.vals, labels=names(mu.vals), pos=4, cex = cex.val)
points(gmin.port$sd, gmin.port$er, pch=16, cex=2.5, col="black")
text(gmin.port$sd, gmin.port$er, labels="Global Min", pos=2, cex = cex.val)
# plot bootstrapped global min
points(sd.gmin.boot, mu.gmin.boot, col="green", pch=16)

# look at bootstrap distribution
par(mfrow=c(2,2))
hist(mu.gmin.boot, col="slateblue1")
qqnorm(mu.gmin.boot, col="slateblue1", pch=16)
qqline(mu.gmin.boot)

hist(sd.gmin.boot, col="slateblue1")
qqnorm(sd.gmin.boot, col="slateblue1", pch=16)
qqline(sd.gmin.boot)
par(mfrow=c(1,1))

bias.mu.gmin = mean(mu.gmin.boot) - gmin.port$er
se.mu.gmin = sd(mu.gmin.boot)
ci.mu.gmin.95 = c(gmin.port$er-2*se.mu.gmin, 
                  gmin.port$er+2*se.mu.gmin)
bias.mu.gmin
se.mu.gmin
ci.mu.gmin.95

bias.sd.gmin = mean(sd.gmin.boot) - gmin.port$sd
se.sd.gmin = sd(sd.gmin.boot)
ci.sd.gmin.95 = c(gmin.port$sd-2*se.sd.gmin, 
                  gmin.port$sd+2*se.sd.gmin)
bias.sd.gmin
se.sd.gmin
ci.sd.gmin.95


# min var portfolio weights
par(mfrow=c(2,2))
hist(w.gmin.boot[, "lvc.z"], main="Large Value", xlab="Weight", col="slateblue1")
hist(w.gmin.boot[, "svc.z"], main="Small Value", xlab="Weight", col="slateblue1")
hist(w.gmin.boot[, "lgc.z"], main="Large Growth", xlab="Weight", col="slateblue1")
hist(w.gmin.boot[, "ec.z"], main="Europe", xlab="Weight", col="slateblue1")
par(mfrow=c(1,1))

# compute bias, mse and 95% CI for weights
bias.w.gmin = colMeans(w.gmin.boot) - gmin.port$weights
se.w.gmin = apply(w.gmin.boot, 2, sd)
ci.w.gmin.95 = rbind(gmin.port$weights-2*se.w.gmin, 
                     gmin.port$weights+2*se.w.gmin)
rownames(ci.w.gmin.95) = c("lower", "upper")
bias.w.gmin
se.w.gmin
ci.w.gmin.95


# sort bootstrap values by mean
tmp.w.boot = w.gmin.boot[1:20, ]
tmp.mu.boot = mu.gmin.boot[1:20, ]
tmp.sd.boot = sd.gmin.boot[1:20, ]
sort.idx = order(tmp.mu.boot)

# look at weights in stacked bar charts
chart.StackedBar(tmp.w.boot[sort.idx,], 
                 xaxis.labels=round(tmp.mu.boot[sort.idx],digits=3), 
                 xlab="Portfolio SD", ylab="Weights", cex.lab=1.5,
                 cex.axis=1.5)

# look at correlation between min var weights
cor(w.gmin.boot)
pairs(w.gmin.boot)

qqnorm(w.gmin.boot[, "svc.z"], col="slateblue1", pch=16)
qqline(w.gmin.boot[, "svc.z"])

# bootstrap efficient frontier
#

# show efficient frontier from sample estimates
ef = efficient.frontier(mu.vals,ecov)

# plot efficient portfolios
plot(ef$sd, ef$er, type="b", ylim=c(0, 0.04), xlim=c(0, 0.17), 
     pch=16, col="blue", cex=2, ylab=expression(mu[p]), xlab=expression(sigma[p]))

points(sd.vals, mu.vals, pch=16, cex=2, col="black")
points(gmin.port$sd, gmin.port$er, pch=16, cex=2, col="green")
text(sd.vals, mu.vals, labels=names(mu.vals), pos=4, cex=2)
text(gmin.port$sd, gmin.port$er, labels="Global min", pos=4, cex=2)

# bootstrap efficient frontier

ef.list = list()
set.seed(123)
for (i in 1:n.boot) {
    boot.idx = sample(n.obs, replace=TRUE)
    ret.boot = a[boot.idx, ] 
    mu.boot = colMeans(ret.boot)
    cov.boot = cov(ret.boot) 
    ef.boot = efficient.frontier(mu.boot, cov.boot)
    ef.list[[i]] = ef.boot
}

# plot sample efficient frontier with first 20 bootstrap 
# efficient frontiers
# plot efficient portfolios
plot(ef$sd, ef$er, type="b", ylim=c(-0.01, 0.03), xlim=c(0.03, 0.07), 
     pch=16, col="blue", cex=0.5, ylab=expression(mu[p]), xlab=expression(sigma[p]))

points(sd.vals, mu.vals, pch=16, cex=0.5, col="black")
text(sd.vals, mu.vals, labels=names(mu.vals), pos=4, cex=0.5)
# plot bootstrap efficient frontiers
for (i in 1:30) {
    points(ef.list[[i]]$sd, ef.list[[i]]$er, type="b",
           pch=16, col=i, cex=0.5)
}

```
